A Kochen-Specker system has at least 22 vectors

### 2014-12-16 paper/main.tex
paper: updates

Before we continue, we would like to make two remarks.
\begin{enumerate}
\item
There is an inconsistency in the literature on how to count the number
of points in a Kochen-Specker system.  One way is to take the minimal
number of points required to be non-010-colorable.
This is how we count in this paper.
Another way is to require that for every pair of orthogonal points
in the system, there is a third point orthogonal to both.
This way of counting is used in~\cite{pavivcic}.
\item
We consider three-dimensional Kochen-Specker systems
for two reasons.
First, every 3-dimensional system can be seen as a higher-dimensional
            system.   The converse if false.
Secondly, the size of the least Kochen-Specker system
            in dimension 4 and higher has already been found.\cite{pavivcic}
            The three-dimensional case is still open.
\end{enumerate}

@@

% attribute mckay for answering questions?
% TODO attribute geng
\clearpage
\bibliography{main}{}
\bibliographystyle{plain}

>>>>

Before we continue, we would like to make two remarks.
\begin{enumerate}
\item
There is an inconsistency in the literature on how to count the number
of points in a Kochen-Specker system.  One way is to take the minimal
number of points required to be non-010-colorable.
This is how we count in this paper.
Another way is to require that for every pair of orthogonal points
in the system, there is a third point orthogonal to both.
This way of counting is for example used in~\cite{pavivcic}.
\item
We consider three-dimensional Kochen-Specker systems
for two reasons.
First, every 3-dimensional system can be seen as a higher-dimensional
            system.   The converse if false.
Secondly, the size of the least Kochen-Specker system
            in dimension 4 and higher has already been found.\cite{pavivcic}
            The three-dimensional case is still open.
\end{enumerate}

@@

\clearpage
\bibliography{main}{}
\bibliographystyle{plain}

\appendix
\clearpage
\section{Appendix}
In this appendix, we show by hand, that indeed the 19 vertex graph
considered at the start of section 3 %TODO
is indeed not embeddable.
We restrict the calculation to the~$13$-vertex unembeddable
subgraph.

\begin{center}
    \footnotesize
    \begin{tikzpicture}[thick,scale=3.5,
        p/.style={fill=white, circle, inner sep=1pt, minimum width=45pt}]
        \draw

                (45:0.9) node[p, align=center]{0 \\ $(1,0,0)$} -- (65:1.3) node[p, align=center]{6 \\$(0,x_3,\cdot)$}
                (45:0.9)  -- (25:1.3) node[p, align=center]{15 \\ $(0,\cdot,\text{-}x_3)$}
                (65:1.3)  -- (25:1.3)

                (90+45:0.9) node[p, align=center]{3 \\ $(\cdot,\text{-}x_1,0)$} -- (90+65:1.3) node[p, align=center]{8 \\ $(g,h,\cdot)$ }
                (90+45:0.9)  -- (90+25:1.3) node[p, align=center]{13 \\ $(e,f,\cdot)$}
                (90+65:1.3)  -- (90+25:1.3)

                (180+45:0.9) node[p, align=center]{10 \\ $(0,1,0)$} -- (180+65:1.3) node[p, align=center]{4 \\ $(x_2,0,\cdot)$}
                (180+45:0.9)  -- (180+25:1.3) node[p, align=center]{14 \\ $(\cdot,0,\text{-}x_2)$}
                (180+65:1.3)  -- (180+25:1.3)

                (270+45:0.9) node[p, align=center]{9 \\$(x_1,\cdot,0)$} -- (270+65:1.3) node[p, align=center]{2 \\ $(a,b,\cdot)$}
                (270+45:0.9)  -- (270+25:1.3) node[p, align=center]{17 \\ $(c,d,\cdot)$}
                (270+65:1.3)  -- (270+25:1.3)

            \foreach \x in {0,180} {
                (\x+25:1.3) -- (\x-25:1.3)
                (\x+65:1.3) -- (\x+115:1.3)
           }

        (90:0.6) -- (90:0.6)  node[p, align=center]{18 \\ $(0,0,1)$}
        (90:0.6) -- (270+45:0.9)
        (90:0.6) -- (270-45:0.9)
        (90:0.6) -- (90+45:0.9)
	(90:0.6) -- (90-45:0.9)

(45:1)--(225:1)
(135:1)--(315:1) ;
    \end{tikzpicture}
\end{center}

This is the subgraph and we have given the vertices names corresponding
to the original, 19 vertex, graph. Below each name is a coordinate
on the sphere (whenever there is a $\cdot$ in the coordinate, we
mean that this component of the coordinate is fixed by the other
components, as it must be a point on the sphere, where we do take
this $\cdot$ to be the positive square root).

Now if two vertices are connected via an edge, we have that the
inner product of their coordinates is zero. Vertices connected to
the points 0,10 and 18 are trivially zero. The non trivial equations
then become:
\begin{align}
&ax_1 + b\sqrt{1-{x_1}^2}=0\\
&cx_1 + d\sqrt{1-{x_1}^2}=0\\
&e\sqrt{1-{x_1}^2} - fx_1=0\\
&g\sqrt{1-{x_1}^2} - hx_1=0\\
&g\sqrt{1-{x_2}^2} - x_2 \sqrt{1-g^2-h^2}=0\\
&cx_2 + \sqrt{1-{x_2}^2}\sqrt{1-c^2-d^2}=0\\
&b\sqrt{1-{x_3}^2} - x_3 \sqrt{1-a^2-b^2}=0\\
&fx_3 + \sqrt{1-e^2-f^2}\sqrt{1-{x_3}^2}=0\\
&ac + bd + \sqrt{1-a^2-b^2}\sqrt{1-c^2 - d^2} = 0\\
&eg + fh + \sqrt{1-e^2-f^2}\sqrt{1-g^2 - h^2} = 0
\end{align}

From the equations containing only $x_1$ we can eliminate four variables.
\begin{equation}
b = \frac{-a x_1}{\sqrt{1-{x_1}^2}} \tag{1'}
\end{equation}
\begin{equation}
d = \frac{-c x_1}{\sqrt{1-{x_1}^2}} \tag{2'}
\end{equation}
\begin{equation}
e = \frac{f x_1}{\sqrt{1-{x_1}^2}} \tag{3'}
\end{equation}
\begin{equation}
g = \frac{h x_1}{\sqrt{1-{x_1}^2}} \tag{4'}
\end{equation}
Note that this is possible since all $x_i \in (-1,1)\backslash
\{0\}$ since otherwise we would have coinciding coordinates of
vertices in the graph.

Substituting these expressions in the other six equations and using
that~$\tfrac{x}{1-x} +1 = \tfrac{1}{1-x}$, we obtain:

\begin{align}
&\frac{h x_1}{\sqrt{1-{x_1}^2}}\sqrt{1-{x_2}^2} - x_2 \sqrt{1-\frac{h^2}{1-{x_1}^2}}=0 \tag{5'}\\
&cx_2 + \sqrt{1-{x_2}^2}\sqrt{1-\frac{c^2}{1-{x_1}^2}}=0 \tag{6'}\\
&\frac{-a x_1}{\sqrt{1-{x_1}^2}}\sqrt{1-{x_3}^2} - x_3 \sqrt{1-\frac{a^2}{1-{x_1}^2}}=0\tag{7'}\\
&fx_3 + \sqrt{1-\frac{f^2}{1-{x_1}^2}}\sqrt{1-{x_3}^2}=0\tag{8'}\\
&ac + \frac{-a x_1}{\sqrt{1-{x_1}^2}}\frac{-c x_1}{\sqrt{1-{x_1}^2}} + \sqrt{1-\frac{a^2}{1-{x_1}^2}}\sqrt{1-\frac{c^2}{1-{x_1}^2}} = 0 \tag{9'}\\
&\frac{f x_1}{\sqrt{1-{x_1}^2}}\frac{h x_1}{\sqrt{1-{x_1}^2}} + fh + \sqrt{1-\frac{f^2}{1-{x_1}^2}}\sqrt{1-\frac{h^2}{1-{x_1}^2}} = 0\tag{10'}
\end{align}

These can then almost trivially be rewritten to:

\begin{align}
&h x_1\sqrt{1-{x_2}^2} - x_2 \sqrt{1-{x_1}^2 - h^2}=0 \tag{5''}\label{5''}\\
&cx_2\sqrt{1-{x_1}^2} + \sqrt{1-{x_2}^2}\sqrt{1-{x_1}^2 - c^2}=0\tag{6''}\label{6''}\\
&a x_1\sqrt{1-{x_3}^2} + x_3 \sqrt{1-{x_1}^2 - a^2}=0\tag{7''}\label{7''}\\
&fx_3\sqrt{1-{x_1}^2} + \sqrt{1-{x_1}^2 - f^2}\sqrt{1-{x_3}^2}=0\tag{8''}\label{8''}\\
&ac + \sqrt{1- {x_1}^2 - a^2}\sqrt{1- {x_1}^2 - c^2} = 0\tag{9''}\label{9''}\\
&fh + \sqrt{1- {x_1}^2 - f^2}\sqrt{1- {x_1}^2 - h^2} = 0\tag{10''}\label{10''}
\end{align}

It then follows from \eqref{7''}, \eqref{9''} and \eqref{6''} that:
\begin{eqnarray*}
\frac{\sqrt{1-{x_3}^2}}{x_3} &=& \frac{-\sqrt{1-{x_1}^2 - a^2}}{ax_1}\\
&=& \frac{c}{x_1 \sqrt{1-{x_1}^2 -c^2}}\\
&=& \frac{-\sqrt{1-{x_2}^2}}{x_1 x_2 \sqrt{1-{x_1}^2}}
\end{eqnarray*}

While using \eqref{5''},\eqref{10''} and \eqref{8''} we find that:
\begin{eqnarray*}
\frac{\sqrt{1-{x_2}^2}}{x_2} &=& \frac{\sqrt{1-{x_1}^2 - h^2}}{hx_1}\\
&=& \frac{-f}{x_1 \sqrt{1-{x_1}^2 -f^2}}\\
&=& \frac{\sqrt{1-{x_3}^2}}{x_1 x_3 \sqrt{1-{x_1}^2}}
\end{eqnarray*}

Inserting this expression for $\frac{\sqrt{1-{x_2}^2}}{x_2}$ into
the expression for $\frac{\sqrt{1-{x_3}^2}}{x_3}$ we obtain the
equation:

\begin{equation}
\frac{\sqrt{1-{x_3}^2}}{x_3} = \frac{-\sqrt{1-{x_3}^2}}{x_3 {x_1}^2  (1-{x_1}^2)}
\end{equation}
But then, eliminating the $x_3$ terms we find:
\begin{equation}
{x_1}^2 (1-{x_1}^2) = -1
\end{equation}
Which is our desired contradicton.
